---
title: "Analysing Data on Barbell Lifts"
output: html_document
---

## Executive Summary

This project develops an algorithm which predicts whether a person does a weight lifting exercise correctly or incorrectly based on data from accelerometers on the belt, forearm, arm, and dumbell of the participants. The data are taken from Groupware@LES's Human Activity Recognition project. 

## Getting and Cleaning Data

We load the data and remove the variables that hypothetically should not have bearing on the prediction algorithm (e.g. the index number, timestamps, window number etc). We also remove variables with a large proportion of missing observations. 

```{r, cache=TRUE}
training <- read.csv("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", na.strings = c("", "#DIV/0!", "NA"))
testing <- read.csv("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", na.strings = c("", "#DIV/0!", "NA"))

library(caret)
                                      
trainreg <- training[,-c(1,3,4,5,6,7, 12:36, 50:59, 69:83, 87:101, 103:112, 125:139, 141:150)]
```

## Creating the Model

We then create a model to predict, based on the data collected, whether the participants executed the exercise correctly (corresponding to Class A) or incorrectly (corresponding to Classes B, C, D and E). 

Using the 'train' function in R's *caret* package, we first pre-process the data using Principal Components Analysis, then create a model using the bootstrap aggregating ('bagging') method. We also use the 'trainControl' function to conduct 40-fold cross-validation. 

The resampling results show in sample accuracy of 97.2%. We therefore expect the out-of-sample error to be larger than 2.8%.

```{r, cache=TRUE}
trctrl <- trainControl(method="cv", number=40)

modFit <- train(trainreg$classe~., method="treebag", preProcess="pca", data=trainreg, trControl=trctrl)

print(modFit)                        
```

 
